Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-10 ICML-submission/from-linear-separability-to-strong-linear-separability.tex
Add margin theorem

We construct the three different mappings $\phi_n, \psi_n, \rho$. The first two
are parameterized by a non-negative integer $n$. The domain of $\phi_n$ and
$\psi_n$ is $\R^d$. The domain of $\rho$ is the unit ball of $\R^d$. The
co-domain of $\phi_n$ and $\psi_n$ is $\R^q$ where $q = \sum_{i=0}^n
\binom{i+d-1}{d-1}$ equipped with the standard inner product $\ip{x}{x'}_{\R_q} =
\sum_{i=1}^p x_i x'_i$. The co-domain of $\rho$ is the classical real separable Hilbert space
$\ell_2 = \{ x \in \R^\infty ~:~ \sum_{i=1}^\infty x_i^2 < + \infty \}$ equipped
with the standard inner product $\ip{x}{x'}_{\ell_2} = \sum_{i=1}^\infty x_i x'_i$. We
index the coordinates of the co-domain $\R^q$ by $d$-tuples $(\alpha_1,
\alpha_2, \dots, \alpha_d)$ of non-negative integers such that $\alpha_1 +
\alpha_2 + \dots + \alpha_d \le n$. Similarly, we index the coordinates of
$\ell_2$ by $d$-tuples $(\alpha_1, \alpha_2, \dots, \alpha_d)$ of non-negative
integers. The mappings are defined by
\begin{align*}
& \left(\phi_n(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)}
= x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d}
\\
& \left(\psi_n(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \\
& \qquad \cdot \sqrt{\binom{n}{\alpha_1 + \alpha_2 + \dots + \alpha_d} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\\
& \left(\rho(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \\
& \qquad \cdot \sqrt{2^{-(\alpha_1 + \alpha_2 + \dots + \alpha_d)} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\end{align*}
where $\binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d} = \frac{(\alpha_1 + \alpha_2 + \dots + \alpha_d)!}{\alpha_1! \alpha_2! \dots \alpha_d!}$
is the multinomial coefficient and $\binom{n}{k} = \frac{n!}{(n-k)!k!}$ is the binomial coefficient.

@@

\begin{lemma}
Let $d \ge 1$ and $n \ge 0$ be integers and let $q = \sum_{i=0}^n \binom{i+d-1}{d-1}$.
Let $p:\R^d \to \R$ be a multivariate polynomial.
\begin{enumerate}
\item If $n \ge \deg(p)$ then there exists $c \in \R^q$ such that $p(x) = \ip{c}{\phi_n(x)}_{\R^q}$
and $\norm{c}_{\R^q} = \norm{p}$.
\item If $n \ge \deg(p)$ then there exists $c' \in \R^q$ such that $p(x) = \ip{c'}{\psi_n(x)}_{\R^q}$
and $\norm{c'}_{\R^q} \le \norm{p}$.
\item There exists $c'' \in \ell_2$ such that $p(x) = \ip{c''}{\psi_n(x)}_{\ell_2}$
and $\norm{c''}_{\R^q} \le 2^{\deg(p)/2} \norm{p}$.
\end{enumerate}
\end{lemma}

@@

>>>>

We construct the three different mappings $\phi_n, \psi_n, \rho$. The first two
are parameterized by a non-negative integer $n$. The domain of $\phi_n$ and
$\psi_n$ is $\R^d$. The domain of $\rho$ is the unit ball of $\R^d$. The
co-domain of $\phi_n$ and $\psi_n$ is $\R^q$ where $q = \sum_{i=0}^n
\binom{i+d-1}{d-1}$ equipped with the standard inner product $\ip{x}{x'}_{\R^q} =
\sum_{i=1}^p x_i x'_i$. The co-domain of $\rho$ is the classical real separable Hilbert space
$\ell_2 = \{ x \in \R^\infty ~:~ \sum_{i=1}^\infty x_i^2 < + \infty \}$ equipped
with the standard inner product $\ip{x}{x'}_{\ell_2} = \sum_{i=1}^\infty x_i x'_i$. We
index the coordinates of the co-domain $\R^q$ by $d$-tuples $(\alpha_1,
\alpha_2, \dots, \alpha_d)$ of non-negative integers such that $\alpha_1 +
\alpha_2 + \dots + \alpha_d \le n$. Similarly, we index the coordinates of
$\ell_2$ by $d$-tuples $(\alpha_1, \alpha_2, \dots, \alpha_d)$ of non-negative
integers. The mappings are defined by
\begin{align*}
& \left(\phi_n(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)}
= x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d}
\\
& \left(\psi_n(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \\
& \qquad \cdot \sqrt{\binom{n}{\alpha_1 + \alpha_2 + \dots + \alpha_d} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\\
& \left(\rho(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \\
& \qquad \cdot \sqrt{2^{-(\alpha_1 + \alpha_2 + \dots + \alpha_d)} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\end{align*}
where $\binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d} = \frac{(\alpha_1 + \alpha_2 + \dots + \alpha_d)!}{\alpha_1! \alpha_2! \dots \alpha_d!}$
is the multinomial coefficient and $\binom{n}{k} = \frac{n!}{(n-k)!k!}$ is the binomial coefficient.

@@

\begin{lemma}[Norm bound]
\label{lemma:norm-bound}
Let $d \ge 1$ and $n \ge 0$ be integers and let $q = \sum_{i=0}^n \binom{i+d-1}{d-1}$.
Let $p:\R^d \to \R$ be a multivariate polynomial.
\begin{enumerate}
\item If $n \ge \deg(p)$ then there exists $c \in \R^q$ such that $p(x) = \ip{c}{\phi_n(x)}_{\R^q}$
and $\norm{c}_{\R^q} = \norm{p}$.
\item If $n \ge \deg(p)$ then there exists $c' \in \R^q$ such that $p(x) = \ip{c'}{\psi_n(x)}_{\R^q}$
and $\norm{c'}_{\R^q} \le \norm{p}$.
\item There exists $c'' \in \ell_2$ such that $p(x) = \ip{c''}{\psi_n(x)}_{\ell_2}$
and $\norm{c''}_{\R^q} \le 2^{\deg(p)/2} \norm{p}$.
\end{enumerate}
\end{lemma}

@@

Using the lemma and the polynomial approximation theorems, we can prove
the mappings $\phi_n, \psi_n, \rho$ map any a weakly separable data
to a strongly separable data set.

\begin{theorem}
Let $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T) \in \R^d \times \{1,2,\dots,K\}$
labeled examples that are weakly linearly separable with margin $\gamma > 0$
such that $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le 1$. Then, the labeled
examples $(\rho(x_1), y_1), (\rho(x_2), y_2), \dots, (\rho(x_T), y_T)$ are
strongly linearly separable with margin
$$
\gamma' = \frac{1}{2\sqrt{K} \left(376 \lceil \log_2(2m) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2m) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}}
$$
and $\norm{\rho(x_1)}_{\ell_2}, \norm{\rho(x_2)}_{\ell_2}, \dots, \norm{\rho(x_T)}_{\ell_2} \le 2$.
TODO
\end{theorem}

\begin{proof}
First osberve that for any $t=1,2,\dots,T$,
$$
\norm{\rho(x_t)}_{\ell_2} = K_{\rho}(x_t,x_t) = \frac{1}{1 - \frac{1}{2} \norm{x_t}^2} \le 2 \; .
$$

Since the examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ are weakly linearly separable
with margin $\gamma$ there are vectors $w_1, w_2, \dots, w_K$ satisfying
\eqref{equation:weak-linear-separability-1} anf \eqref{equation:weak-linear-separability-2}.

Fix any $i \in \{1,2,\dots,K\}$. Consider the $K-1$ vectors
$(w_i - w_j)/2$ for $j \in \{1,2,\dots,K\} \setminus \{i\}$. Note that the vectors
have norm at most $1$.
\autoref{theorem:polynomial-approximation-1} implies that
there exists a polynomial $p_i:\R^d \to \R$ such
that
$$
\deg(p) = \lceil \log_2(2m) \rceil \cdot \lceil \sqrt{\frac{2}{\gamma}} \rceil \; ,
$$
for all $t=1,2,\dots,T$, if $y_t = i$ then $p_i(x_t) \ge 1/2$ and if $y_t \neq i$ then $p_i(x_t) \le -1/2$,
and
$$
\norm{p} \le \left(188 \lceil \log_2(2m) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2m) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \; .
$$
By \autoref{lemma:norm-bound} there exists $c_i \in \ell_2$ such that
$\ip{c_i}{\rho(x)} = p(x)$ and
$$
\norm{c_i}_{\ell_2}
\le \left(376 \lceil \log_2(2m) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2m) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \; .
$$
Define vector $u_i \in \ell_2$ as
$$
u_i = \frac{c_i}{\sqrt{K} \left(376 \lceil \log_2(2m) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2m) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}} \; .
$$

Then, $\norm{u_1}^2 + \norm{u_2}^2 + \dots + \norm{u_K}^2 \le 1$ and
for all $t=1,2,\dots,T$, $\ip{u_{y_t}}{x_t} \ge \gamma'$
and for all $j \in \{1,2,\dots,K\} \setminus \{y_t\}$,
$\ip{u_j}{x_t} \le - \gamma'$. In other words,
$(\rho(x_1), y_1), (\rho(x_2), y_2), \dots, (\rho(x_T), y_T)$ are
strongly linearly separable with margin $\gamma'$.
\end{proof}

### 2019-01-10 ICML-submission/related-work.tex
Add margin theorem

The problem of bandit multiclass linear classification was initially formulated
by the pioneering work of~\citet{Kakade-Shalev-Shwartz-Tewari-2008}. In that
work, two computationally inefficient algorithms that work in the standard
linearly separable setting are proposed: one with a mistake bound of $O(K^2 d
\ln(\frac{d}{\gamma^2}))$, the other with a mistake bound of
$\widetilde{O}(\frac{K^2}{\gamma^2} \ln T)$. In addition, the authors propose
the \textsc{Banditron} algorithm, a computationally efficient algorithm that has
a $O(T^{2/3})$ regret against the multiclass hinge loss in the general setting,
and has a $O(\sqrt{T})$ mistake bound in the $\gamma$-weak linearly separable
setting. The polynomial dependence of \textsc{Banditron}'s mistake bound on the
time horizon is undesirable for problems with a long time horizon. One key open
question left by~\citet{Kakade-Shalev-Shwartz-Tewari-2008} is whether one can
design computationally efficient algorithms that achieve mistake bounds that
match or improve over those of inefficient algorithms. In this paper, we take a
step towards answering this question, showing that efficient algorithms with
mistake bounds quasipolynomial in the margin parameter can be obtained.

@@

\citet{Chen-Lin-Lu-2014} and \citet{Zhang-Jung-Tewari-2018} study online bandit
multiclass boosting under bandit feedback, where one can view online boosting as
online linear classification by treating each base hypothesis as a separate
feature. \citet{Chen-Lin-Lu-2014}'s online weak learning condition implies that
the set of examples is strongly separable by a convex combination of base
hypotheses with a margin. Under this condition, it gives an algorithm with a
$O(T^{3/4})$ mistake bound. \citet{Zhang-Jung-Tewari-2018} considers an online
weak learning condition, which implies that all examples is separable by a
convex combination of base hypotheses with a margin
(See~\cite{Mukherjee-Schapire-2013}, Theorem 3). Under this condition, it gives
a boosting algorithm with a $O(\sqrt{T})$ mistake bound with the knowledge of
the edge parameter. In addition, it gives an adaptive algorithm with a
$O(T^{3/4})$ mistake bound.

@@

There has been much literature on the contextual bandit learning
problem~\citep{Auer-2003, Langford-Zhang-2008}. In this problem, the learner is
given a policy class $\Pi$, and at each timestep $t$, an example $x_t$ is shown,
with an associated cost vector $c_t$ hidden. The learner then takes an action
$\hat{y}_t$, and observed the incurred cost $c_t$. The goal of the learner is to
minimize its regret $\sum_{t=1}^T c_t(\hat{y}_t) - \min_{\pi \in \Pi}
\sum_{t=1}^T c_t(\pi(x_t))$. The online bandit multiclass learning problem is a
special case of the contextual bandits problem, where the cost $c_t(i)$ is the
classification error $I(i \neq y_t)$, and the policy class $\Pi$ is the set of
linear classifiers $\cbr{x \mapsto \argmax_y (Wx)_y: W \in \R^{k \times d}}$. A
series of works in the contextual bandits learning literature focus on
oracle-efficiency~\citep{Dudik-Hsu-Kale-Karampatziakis-Langford-Reyzin-Zhang-2011,
Agarwal-Hsu-Kale-Langford-Li-Schapire-2014, Rakhlin-Sridharan-2016,
Syrgkanis-Krishnamurthy-Schapire-2016,
Syrgkanis-Luo-Krishnamurthy-Schapire-2016}. Specifically, they assume access to
a policy optimization oracle that receives a set of cost-sensitive learning
examples and policy class $\Pi$, return the policy in $\Pi$ that minimizes its
empirical cost on the examples, and the algorithm only call the oracle a
polynomial number of times. However, these algorithms are not truly
computationally efficient in our setting, as it is NP-hard in general to find a
linear classifier that minimizes the empirical cost over a set of
examples~\citep{Arora-Babai-Stern-Sweedyk-1997}.

>>>>

The problem of bandit multiclass linear classification was initially formulated
in the pioneering work of~\citet{Kakade-Shalev-Shwartz-Tewari-2008}. They proposed
two computationally inefficient algorithms that work in the standard
linearly separable setting, one with a mistake bound of $O(K^2 d
\ln(\frac{d}{\gamma^2}))$, the other with a mistake bound of
$\widetilde{O}(\frac{K^2}{\gamma^2} \ln T)$. In addition, the authors propose
the \textsc{Banditron} algorithm, a computationally efficient algorithm that has
a $O(T^{2/3})$ regret against the multiclass hinge loss in the general setting,
and has a $O(\sqrt{T})$ mistake bound in the $\gamma$-weak linearly separable
setting. The polynomial dependence of \textsc{Banditron}'s mistake bound on the
time horizon is undesirable for problems with a long time horizon. One key open
question left by~\citet{Kakade-Shalev-Shwartz-Tewari-2008} is whether one can
design computationally efficient algorithms that achieve mistake bounds that
match or improve over those of inefficient algorithms. In this paper, we take a
step towards answering this question, showing that efficient algorithms with
mistake bounds quasipolynomial in the margin parameter can be obtained.

@@

\citet{Chen-Lin-Lu-2014} and \citet{Zhang-Jung-Tewari-2018} study online bandit
multiclass boosting under bandit feedback, where one can view online boosting as
online linear classification by treating each base hypothesis as a separate
feature. \citet{Chen-Lin-Lu-2014}'s online weak learning condition implies that
the set of examples is strongly separable by a convex combination of base
hypotheses with a margin. Under this condition, it gives an algorithm with a
$O(T^{3/4})$ mistake bound. \citet{Zhang-Jung-Tewari-2018} considers an online
weak learning condition, which implies that all examples is separable by a
convex combination of base hypotheses with a margin; see~\citet[Theorem
3]{Mukherjee-Schapire-2013}. Under this condition, it gives a boosting algorithm
with a $O(\sqrt{T})$ mistake bound with the knowledge of the edge parameter. In
addition, it gives an adaptive algorithm with a $O(T^{3/4})$ mistake bound.

@@

There has been much literature on the contextual bandit learning
problem~\citep{Auer-2003, Langford-Zhang-2008}. In this problem, the learner is
given a policy class $\Pi$, and at each timestep $t$, an example $x_t$ is shown,
with an associated cost vector $c_t$ hidden. The learner then takes an action
$\widehat{y}_t$, and observed the incurred cost $c_t$. The goal of the learner is to
minimize its regret $\sum_{t=1}^T c_t(\widehat{y}_t) - \min_{\pi \in \Pi}
\sum_{t=1}^T c_t(\pi(x_t))$. The online bandit multiclass learning problem is a
special case of the contextual bandits problem, where the cost $c_t(i)$ is the
classification error $\indicator{i \neq y_t}$, and the policy class $\Pi$ is the set of
linear classifiers $\cbr{x \mapsto \argmax_y (Wx)_y: W \in \R^{k \times d}}$. A
series of works in the contextual bandits learning literature focus on
oracle-efficiency~\citep{Dudik-Hsu-Kale-Karampatziakis-Langford-Reyzin-Zhang-2011,
Agarwal-Hsu-Kale-Langford-Li-Schapire-2014, Rakhlin-Sridharan-2016,
Syrgkanis-Krishnamurthy-Schapire-2016,
Syrgkanis-Luo-Krishnamurthy-Schapire-2016}. Specifically, they assume access to
a policy optimization oracle that receives a set of cost-sensitive learning
examples and policy class $\Pi$, return the policy in $\Pi$ that minimizes its
empirical cost on the examples, and the algorithm only call the oracle a
polynomial number of times. However, these algorithms are not truly
computationally efficient in our setting, as it is NP-hard in general to find a
linear classifier that minimizes the empirical cost over a set of
examples~\citep{Arora-Babai-Stern-Sweedyk-1997}.

### 2019-01-10 ICML-submission/title-and-abstract.tex
Rename linear separability to weak linear separability
Add margin theorem

\begin{abstract}
We study the problem of efficient online multiclass linear classification with
bandit feedback, where all examples belong to one of $K$ classes and lie in the
$d$-dimensional Euclidean space. Previous works have left open the challenge of
designing efficient algorithms with finite mistake bounds when the data is
linearly separable by a margin $\gamma$. In this work, we take a first step
towards this problem. We consider two notions of linear separability: strong
linear separability and weak linear separability.

@@

\item Under the more challenging weak linear separability condition, we design
an efficient algorithm with a mistake bound of $2^{\widetilde{O}(\min(K \log^2
\frac{1}{\gamma}, \log K \sqrt{\frac 1 \gamma}))}$ \footnote{We use the notation
$\widetilde{O}(f(\cdot)) = O(f(\cdot) \polylog(f(\cdot)))$.}. Our algorithm is
based on a infinite-dimensional feature mapping that transforms weak linear
separable examples to strong linear separable examples, which in turn is
inspired by \cite{Klivans-Servedio-2008}'s work on learning intersection of
halfspaces.
\end{enumerate}
\end{abstract}

>>>>

\begin{abstract}
We study the problem of efficient online multiclass linear classification with
bandit feedback, where all examples belong to one of $K$ classes and lie in the
$d$-dimensional Euclidean space. Previous works have left open the challenge of
designing efficient algorithms with finite mistake bounds when the data is
linearly separable by a margin $\gamma$. In this work, we take a first step
towards this problem. We consider two notions of linear separability,
\emph{strong} and \emph{weak}.

@@

\item Under the more challenging weak linear separability condition, we design
an efficient algorithm with a mistake bound of $2^{\widetilde{O}(\min(K \log^2
\frac{1}{\gamma}, \log K \sqrt{\frac 1 \gamma}))}$ \footnote{We use the notation
$\widetilde{O}(f(\cdot)) = O(f(\cdot) \polylog(f(\cdot)))$.}. Our algorithm is
based on a infinite-dimensional feature mapping that transforms weak linear
separable examples to strong linear separable examples, which in turn is
inspired by the work of \citet{Klivans-Servedio-2008} on learning intersection of
halfspaces.
\end{enumerate}
\end{abstract}
